---
title: "Code"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("corrplot")
install.packages("car")
install.packages("shiny")
install.packages("shinydashboard")
install.packages("MASS")
```

```{r}
library("tidyverse")
library("shiny")
library("shinydashboard")
library("car")
library("corrplot")
library("ggplot2")
library("MASS")
```

## Reading dataset
```{r}
library(readr)
tracks_features <- read_csv("archive/tracks_features.csv")
```

## Adding data into a dataframe
```{r}
musicData <- data.frame(tracks_features)
```

## Using random sampling to select 5000 observations from the dataset for computational limitations
```{r}
library(dplyr)
set.seed(12345)  
musicData <- sample_n(musicData, 5000, replace = FALSE, prob = NULL)
musicData
```

## Observing variable information
```{r}
library("tidyverse")
str(musicData)
```


## Data exploration, univariate box plots
```{r}
boxplot(musicData$danceability)
boxplot(musicData$energy)
boxplot(musicData$key)
boxplot(musicData$loudness)
boxplot(musicData$speechiness)
boxplot(musicData$acousticness)
boxplot(musicData$instrumentalness)
boxplot(musicData$liveness)
boxplot(musicData$valence)
boxplot(musicData$tempo)

```

## Univariate histograms to track distributions
```{r}
hist(musicData$danceability)
hist(musicData$energy)
hist(musicData$key)
hist(musicData$loudness)
hist(musicData$speechiness)
hist(musicData$acousticness)
hist(musicData$instrumentalness)
hist(musicData$liveness)
hist(musicData$valence)
hist(musicData$tempo)
```


## Covariance matrix 
```{r}
library("tidyverse")
numeric_features <- sapply(musicData, is.numeric)
music_num <- musicData[numeric_features]
cov(music_num)
```

## Correlation matrix
```{r}
library("tidyverse")
numeric_features <- sapply(musicData, is.numeric)
music_num <- musicData[numeric_features]
cors <- cor(music_num)
corrplot(cors, method = "number")
```

## Discriminant analysis 
```{r}
music_num <- music_num[ , c("key",    # Reorder data frame
                       names(music_num)[names(music_num) != "key"])]
                     # Print updated data frame
music_num <- music_num[ -c(2, 3, 14, 15, 16) ]
music_num
```

```{r}
music_num[, -1] <- scale(music_num[, -1])
music_num
```

```{r}
apply(music_num[, -1], 2, mean)
```

```{r}
apply(music_num[, -1], 2, sd)
```

## Train Test Split
```{r}
set.seed(1)

# 70% for training 30% for test 
sample <- sample(c(TRUE, FALSE), nrow(music_num), replace = TRUE, prob = c(0.7, 0.3))
train <- music_num[sample, ]
test <- music_num[!sample, ]
train
test

```

## Linear Discriminant Analysis
```{r}
model <- lda(train$key~., data = train)
model
```

```{r}
predicted <- predict(model, test)
names(predicted)
```
```{r}
head(predicted$class)
```
```{r}
head(predicted$posterior)
```
```{r}
head(predicted$x)
```

```{r}
mean(predicted$class==test$key)
```

## Hierarchical clustering

```{r}
install.packages("factoextra") 
install.packages("FactoMineR")
```


```{r}
library(factoextra)
library(FactoMineR)
library(cluster)
```



```{r}
music_num_for_clustering <- music_num

key_label <- music_num_for_clustering$key

music_num_for_clustering$key <- NULL
music_num_for_clustering
```



```{r}
# Dissimilarity matrix
d <- dist(music_num_for_clustering, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
```

## K-means clustering 

```{r}
fviz_nbclust(music_num_for_clustering, kmeans, method = "wss")
```
```{r}
set.seed(1)

km <- kmeans(music_num_for_clustering, centers = 5, nstart = 25)
km
```

```{r}
fviz_cluster(km, data = music_num_for_clustering)
```
```{r}
km$iter
```


```{r}
install.packages("psych")
```
```{r}
km.res <- eclust(music_num_for_clustering, k = 5, nstart=25, graph = FALSE)
```


```{r}
fviz_silhouette(km.res, palette='jco', ggtheme = theme_classic())

```

 
 
 
## Hotelling's T2 test for categorical variable explicit

```{r}
set.seed(10)
hypothesisData <- musicData[ ,c(9:13, 15:20)]
hypothesisData <- sample_n(hypothesisData, 5000, replace = FALSE, prob = NULL)
explicitLabel <- hypothesisData$explicit
hypothesisData$explicit <- NULL
hypothesisData
 
```
```{r}
install.packages("mvnormtest")
```


```{r}
library(mvnormtest)
```


## Testing normality using mshapiro test
```{r}
mshapiro.test(t(hypothesisData))
```

```{r}
det(cov(hypothesisData))
```

```{r}
install.packages('ICSNP')
```


```{r}
library(ICSNP)
```

```{r}
HotellingsT2(filter(hypothesisData, explicitLabel == "TRUE")[, 1:10],
             filter(hypothesisData, explicitLabel == "FALSE")[, 1:10])
```

## PCA
```{r}
music_num_for_pca <- music_num

library(tidyverse)
pca_results <- prcomp(music_num_for_pca, scale = T)
```

```{r}
pca_results$rotation <- -1*pca_results$rotation
pca_results$rotation
```
```{r}
pca_results$sdev / sum(pca_results$sdev^2)
```

```{r}
var_explained <- pca_results$sdev^2 / sum(pca_results$sdev^2)
```

```{r}
library(ggplot2)

qplot(1:11, var_explained) +
  geom_line() +
  geom_point(size=4)+
  xlab("Principal Component") +
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0, 1)
```


## MDS
```{r}
mds <- cmdscale(d, eig = T, k=2)
```
```{r}
mds$GOF
```
```{r}
mds
```


## Canonical correlation


```{r}
dataset1_cancor <- music_num[ which(music_num$danceability>0.036), ]
dataset2_cancor <- music_num[ which(music_num$danceability<=0.036), ]
dataset2_cancor <- dataset2_cancor[-c(0:8), ]
```

```{r}
dataset1_cancor
dataset2_cancor
```

```{r}
install.packages("CCA")

```
```{r}
library(CCA)
cc <- matcor(dataset1_cancor, dataset2_cancor)
```

```{r}
cc$XYcor
```

```{r}
cc1 <- cc(dataset1_cancor, dataset2_cancor)
cc1$cor
```

```{r}
canoncial_loadings <- comput(dataset1_cancor, dataset2_cancor, cc1)
canoncial_loadings
```

```{r}
cc_results$cor
```

## Outlier Detection
```{r}
# Scatter matrix
pairs(music_num)
```

```{r}
mahalanobis(music_num, colMeans(music_num), cov(music_num))

```
```{r}
music_num$outlier <- "No"
music_num$outlier[music_num$p < 0.05] <- "Yes" 
```

```{r}
music_num$outlier[music_num$p < 0.05]
```

```{r}
# New dataset without outliers
music_num_no_outliers <-music_num[!(music_num$outlier == "Yes"),]
music_num_no_outliers <- music_num_no_outliers[-c(12, 13, 14)]
```

## Revisiting techniques after outlier removal 

```{r}
music_num_no_outliers
```

## Same Techniques after removing outliers
```{r}
# correlation
cors_2 <- cor(music_num_no_outliers)
corrplot(cors_2, method = "number")
```

```{r}
# covariance
cov(music_num_no_outliers)
```

```{r}
# LDA
music_num_no_outliers[, -1] <- scale(music_num_no_outliers[, -1])
music_num_no_outliers
apply(music_num_no_outliers[, -1], 2, mean)
apply(music_num_no_outliers[, -1], 2, sd)
set.seed(1)

# 70% for training 30% for test 
sample2 <- sample(c(TRUE, FALSE), nrow(music_num_no_outliers), replace = TRUE, prob = c(0.7, 0.3))
train2 <- music_num_no_outliers[sample, ]
test2 <- music_num_no_outliers[!sample, ]
train2
test2

```
```{r}
# LDA Model 2
model2 <- lda(train2$key~., data = train2)
model2
```

```{r}
predicted <- predict(model, test)
names(predicted)
mean(predicted$class==test$key)

```

```{r}
# hc clustering
music_num_for_clustering2 <- music_num_no_outliers

key_label2 <- music_num_for_clustering2$key

music_num_for_clustering2$key <- NULL
music_num_for_clustering2
```

```{r}
# Dissimilarity matrix
d2 <- dist(music_num_for_clustering2, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc2 <- hclust(d2, method = "complete" )

# Plot the obtained dendrogram
plot(hc2, cex = 0.6, hang = -1)
```

```{r}
# k-means
fviz_nbclust(music_num_for_clustering2, kmeans, method = "wss")
```

```{r}
set.seed(1)

km2 <- kmeans(music_num_for_clustering2, centers = 6, nstart = 25)
km2
```

```{r}
fviz_cluster(km2, data = music_num_for_clustering2)
```

```{r}
# hotelling test
set.seed(10)
hypothesisData2 <- sample_n(music_num_for_clustering2, 4360, replace = FALSE, prob = NULL)
explicitLabel <- hypothesisData2$explicit
hypothesisData2

HotellingsT2(filter(hypothesisData2, explicitLabel == "TRUE")[, 1:10],
             filter(hypothesisData2, explicitLabel == "FALSE")[, 1:10])
```

```{r}
# pca
pca_results2 <- prcomp(music_num_no_outliers, scale = T)
pca_results2$sdev / sum(pca_results2$sdev^2)
```

```{r}
var_explained2 <- pca_results2$sdev^2 / sum(pca_results2$sdev^2)
library(ggplot2)

qplot(1:11, var_explained2) +
  geom_line() +
  geom_point(size=4)+
  xlab("Principal Component") +
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0, 1)
```

```{r}
mds2 <- cmdscale(d2, eig = T, k=2)
mds2$GOF
```

```{r}
# cancor
dataset1_cancor2 <- music_num_no_outliers[ which(music_num_no_outliers$danceability>-0.05), ]
dataset2_cancor2 <- music_num_no_outliers[ which(music_num_no_outliers$danceability<=-0.05), ]
dataset2_cancor2 <- dataset2_cancor[-c(0:51), ]

```

```{r}
cc2 <- matcor(dataset1_cancor2, dataset2_cancor2)
cc2 <- cc(dataset1_cancor2, dataset2_cancor2)
cc2$cor
```
```{r}
canoncial_loadings2 <- comput(dataset1_cancor2, dataset2_cancor2, cc2)
canoncial_loadings2
```
```{r}
head(musicData)
```

```{r}
chisq.test(musicData$artists, musicData$explicit)
```


